---
title: "CF_Flowchart"
author: "Sooyeong Lim"
date: "April 24, 2019"
output: html_document
---
# Step by step illustration of data preprocess procedure in CF data


```{r setup, include=FALSE}
library(dplyr)
library(ggplot2)
library(DiagrammeR)
rm(list=ls())
load("C:/Users/LIMK2Z/Desktop/jittered_tobi.RData")


```




```{r flowchart, echo=FALSE}
grViz("digraph flowchart {
      # node definitions with substituted label text
      node [fontname = Helvetica, shape = rectangle]        
      tab1 [label = '@@1']
      tab2 [label = '@@2']
      tab3 [label = '@@3']
      tab4 [label = '@@4']
      tab5 [label = '@@5']
      tab6 [label = 'Patients who initially diagnosed chronic p.a infection at age over 21 (n=8,066) \n Patients  who diagnosis chornic p.a infection before age 21 but sum p.a is less than 4(n=14,961)']
      tab7 [label = 'Patients whose follow up visit is beyond 6-12 month']
      # edge definitions with the node IDs
      tab1 -> tab2 -> tab3 -> tab4 
      tab1->tab5
      tab2->tab6;


      }

      [1]: 'Cystic fibrosis registrar data (n=43,038)'
      [2]: 'Patients diagnosed as chronic p.a infection (n=38,615)'
      [3]: 'Patients with chronic p.a infection & cumulative p.a sum >=4 at age less than 21 (n=15,588)'
      [4]: 'Patients whose follow up visit is within 6-12 month from baseline visit (n=13,720)'
      [5]: 'Patients who are not diagnoed chronic p.a infection (n=4,423)'
      ")




```

### 1. Decide Subjects with Chronic P.A Infection

* In the original jittered data, there are 43,038 subjects
* In this study, we are interested in subjects with chronic p.a infection, and we defined chronic p.a subjects with subjects whose bacterialculturedone>=4. 38615 subjects are considered as chronic P.A infection.

```{r step2}
## Convert na's in tobi, pa, bacterialculturedone as 0
jittered_data$bacterialculturedone[is.na(jittered_data$bacterialculturedone)]<-0
jittered_data$tobi[is.na(jittered_data$tobi)]<-0
jittered_data$pa[is.na(jittered_data$pa)]<-0

## Define cumulative sum of bacterial culturedone
step1<-jittered_data %>% 
  group_by(eDWID) %>% 
  mutate(cum_bact=cumsum(bacterialculturedone),
                         cum_tobi=cumsum(tobi),
                         cum_pa=cumsum(pa))
# Decide whether the subjects are chornic p.a or not


step2<-step1 %>% mutate(culture_pa_gt_4=case_when(cum_bact>=4 & cum_pa>=4~1,
                                             cum_bact==4 & cum_pa<4~0,
                                             cum_bact<4 ~0),
                        chronic_pa=case_when(cum_bact>=4 ~1,
                                             cum_bact<4 ~0)
                        )
#Print out sample observations
head((step2 %>% select(eDWID, encounterage, chronic_pa,cum_bact,tobi,cum_pa,culture_pa_gt_4)),10)
# Sort out patients who was never treated as chronic p.a
# There are 38,615 subjects who was dianosied as chrnoic p.a
Id_chronic<-step2 %>% group_by(eDWID) %>% filter(row_number()==n() & culture_pa_gt_4==1) %>% select(eDWID)

# count_chronic<-step2 %>% group_by(eDWID) %>% filter(row_number()==n() & chronic_pa==1) %>% select(eDWID)
# nrow(data.frame(unique(count_chronic$eDWID)))
# Filter those subjects who was not diagnosied as Chornic p.a & cumsum(pa)<4
step3<-inner_join(step2, Id_chronic)

```

### 2. Set Baseline


* To decide baseline, bacterial culture done>=4 & cumulative pa sum>=4 is required

* There are 15,588 subjects satisfy 6-21 y.o baseline criterion

```{r step_baseline}
### Select subjects with chroinc_pa==1 & 6<=encounterage<=21
baseline_step1<-step3 %>% filter(encounterage>=6 & encounterage<=21 & !is.na(FEV1) & culture_pa_gt_4==1)

### FFirst observation from & FEV1 !=NA 
baseline_step2<-baseline_step1 %>% group_by(eDWID) %>% select(eDWID,FEV1,tobi,encounterage, cum_tobi, culture_pa_gt_4) 
Id_baseline<-baseline_step2 %>% group_by(eDWID) %>% 
  filter(row_number()==1) %>% mutate(baseline=1) %>% select(eDWID,encounterage, FEV1,baseline)



## There are 15,588 subjects which was diagnoised as chronic p.a before 21 yr old.
nrow(Id_baseline)

## Assign baseline
step4<-left_join(step3,Id_baseline,by=c('eDWID','encounterage','FEV1'))

### Select subjects who only have a baseline (6 to 21 encounterage)
step4<-step4 %>% filter(eDWID %in% Id_baseline$eDWID)

### How many subjects don't have a baseline & sort
base_FEV1<-step4 %>%  filter(baseline==1) %>% group_by(eDWID) %>%
summarize(baseFEV1=FEV1,baseage=encounterage)
step5<-left_join(step4,base_FEV1)

```


### 3. Set Endpoint

* Assign endpoint with patients who did follow up visit with in 6 to 12 month timeframe

* Remove patients who didn't show up after baseline
```{r step_endpoint}
### Decide endpoint
# Select endpoint btw 6~12 month later from baseline
end_step1<-step5 %>% group_by(eDWID) %>% 
  filter(!is.na(FEV1)& encounterage>=baseage+0.5 & encounterage<=baseage+1)
end_step2<-end_step1 %>% group_by(eDWID) %>% filter(row_number()==1) %>% summarize(endpoint=1, endpointFEV1=FEV1, endpointage=encounterage)

step6<-left_join(step5,end_step2)
# Filter out subjects with no endpoint  
step7<-step6 %>% filter(!is.na(endpointage))
## There are 13,720 subects for analysis

## Since we have baseline & endpoint's FEV1, encounterage drop indicator
var<-names(step7) %in% c('baseline','endpoint')
step7<-step7[!var]




save(step7, file="pre_process.Rdata")


```



